### 2020-03-23 sections/discussion.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

% !TEX root = ../main.tex
This paper presents a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling. We introduce an adaptation of the Anderson-Darling test to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.

The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis, which assumes all arms in a given mode have the same pitch angle.

Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for 99\% of the MCMC draws. As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.

We we do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.

As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative. However, we believe that the methodlogy proposed here is a scalable, robust solution to the problems facing investigation of spiral morphology.