### 2020-06-29 sections/abstract.tex
final proofread before submission to MNRAS

% !TEX root = ../main.tex
Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution. In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project \citep{2020arXiv200610450L}. This approach better deals with the large variations in pitch angle between arms in a galaxy than existing methods, and obtains full posterior distributions on parameters. We utilise fitted pitch angles to examine links between bulge and bar strength and pitch angle reported in other work, finding no correlation in our data. We similarly test the simple model of spiral winding proposed by \citet{2019arXiv190910291P}, which assumes the uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and recurrent spiral pitch angle.