### 2021-01-19 referee_response.md
Finished edits to intro

# Referee Response

The authors thank the referee for a careful reading, with useful suggestions for additional citations. We respond in line below to specific points. Major additions to the text are hi-lighted in bold, while major removals are shown with strike-through.

> Assistant Editor's Comments:

> Please remove the references from your abstract.

This has been done.

> Reviewer's Comments:

> This paper uses data from the Galaxy Zoo project to address the problem of the pitch angle of spiral arms in disc galaxies. It uses citizen science data products of the Galaxy Builder project, in particular those on the pitch angle of spiral galaxies. Interesting results are obtained from data supplied by volunteers, which eventually should be published. However, the paper is not yet very satisfactory, and I recommend that a major revision should be made. I am willing to see the revised version again.

We are pleased the referee thinks our results are interesting and should be published.

> 1. Introduction

> The introduction shows already some problems: in Figure 1 examples are shown of grand design spirals, multi-arm spirals and flocculent spirals. However, it is easy to show that the spirals shown at the right column of Figure 1 are not flocculent, by rotating their image by 180°, and overlaying them on the original with 50% transparency: this shows that the top-right spiral has two arms in the central parts, and is multi-armed in the outer parts, while the bottom-right image is that of a barred spiral with predominantly two arms.

The figure and caption has been updated.

> Another problem concerns the Hubble classification, which for spirals is aptly summarized by Sandage (2005), who writes "It was simplicity itself with wide classification bins that merged seamlessly into one another according to three criteria of size of the central bulge, the degree of resolution into condensations (in the language of Reynolds 1920b, 1927a,b), and the openness of the spiral arms." It is very problematic that the authors think they can replace the criterion of "the degree of resolution into condensations" by "how obvious spiral patterns are". Sandage was consistent in his write-ups, and the same criterion of the degree of resolution can be found in his 1961 "Hubble Atlas of Galaxies". He mentions in his 2005 paper that the Elmegreen & Elmegreen spiral arm classes are "important for the physics, but need not be made general in the [classification] notation".

We have scaled down the mention of Elmegreen & Elmegreen spiral classes, and now use the traditional language of "degree of resolution" in the description the Hubble spiral sequence.

> OK, this was the 1st paragraph in the Introduction. The second paragraph is hardly better: A first sentence discusses "a majority of the population of young stars ..." but switches subject to spiral arms: they may trigger star formation, though their main role maybe sweeping up material ... The second sentence asserts that "this rearrangement of) disc gas can lead to the formation of disc-like bulges (... pseudobulges ...), which are prevalent in most spiral galaxies (Kruk et al. 2018). However, in the latter paper this rearrangement of gas is attributed to the bar, and not to the spiral arms…

There is evidence that both bars and spirals do rearrange disc material - this paragraph is intended to act as a short summary of the impact spirals have on galaxies. It has been reworded.

> Let me point out that the motivation for the Lin & Shu (1964) paper was to overcome the 'winding dilemma' posed by differentially rotating spirals such as M31 and NGC 5055 (Prendergast & Burbidge 1960, Oort 1962), and that the statement of preferentially m = 2 armed spirals is in direct contradiction with the numbers given in Figure 1.

We have removed both the mention of m=2, and the numbers from Figure 1. Neither of these are central to the point of our paper.

> Furthermore, the mechanisms of wave amplification are not the same as those behind the assumption of quasi stationary density wave theory. The term "swing amplification" was coined by Toomre (1981), even though he relied on the Goldreich and Lynden-Bell (1965) work.

We have added a clarification to this point.

> What to make of the sentence "Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms" (p3, left column, line 47). They don't correlate anymore?

This was a grammatical error. Thank you for bring it to our attention. It has been fixed.

> The next paragraph then jumps to "Section 3.2 examines ...", and "Section 3.3 investigates ...", without explaining what is to come in Section 2 and 3.1…

We have fixed this omission.

> Section 2.1

> The authors say: "We assume that a galaxy has some value for pitch angle, \phi_{gal}, and that the pitch angles of spiral arms in that galaxy, \phi_{arm},
are constant with radius (giving logarithmic spirals)" This is very unclear. If the pitch angle is constant with radius, then in all the segments \phi_{i} in eq. (2) will be the same and \phi_{gal}=\phi_{arm}. So why introduce \phi_{gal}? Why introduce two different definitions of the pitch angle, if there is only one value? How does this hypothesis influence the results in this paper?

We have added text to make it clearer that different arms can have different pitch angles, even if in a single arm pitch angle may be constant with radius.

> On page 3 it is written: "A common assumption  when  measuring  galaxy pitch angle is that observed spiral arms have a constant pitch angle with radius (e.g. Davis et al. 2012;.... " This is repated a number of times in other parts of the paper. Yet this is wrong, as first duscussed by Considere and Athanassoula (who introduced the 2DFFT method) and as stressed further in Davis et al 2012.

Davis et al. 2012 say that the inclusion of non-logarithmic spirals (with kinks) can negatively impact measurements using 2D FFT.

Davis et al say in their paper: "As pointed out by Considere &  Athanassoula
(1988), this  method does  not  assume  that observed spiral structures are
logarithmic. It only decomposes the observed distributions into a superposition
of logarithmic spirals of different pitch angles and number of arms, which can
be thought of as building blocks." The authors should read the Considere &
Athanassoula paper carefully, as well as the Davis et al (2012) paper and change
accordingly all parts of the paper where they make use of this assumption, or
find a different way to justify it.

The first line of the abstract of Davis et al (2012) is *"A logarithmic spiral is a prominent feature appearing in a majority of observed galaxies"*. What more justification for their use would you want?

Some of the assumptions made in this paper come without much/any justification.
The authors should discuss the effects of each of the assumptions for which they
offer no justification. E.g. is it reasonable to assume that the spread sigma_[gal]
is constant? What do results from 2DFFT give? Some *quantitative* analysis of
results from previous work would be useful here. If this assumption is simply
based on lack of any better knowledge, the effect of this assumption should be
tested numerically and discussed. But the authors could first check the literature
whether there is better knowledge from previous 2DFFT work.

An ANTI-TODO. I don’t think we should do this (apply P2DFFT). “this is definitely valid, and wasn't done originally due to time constraints. I'm not sure how easy it is to apply P2DFFT, but will investigate:

Section 2.2

This section is rather sparse in information. However, there are 198 galaxies
considered, of which 98 galaxies were shown twice to volunteers. In Lingard
et al. 2020, the Galaxy Builder paper - GBpaper for short - the stellar masses
are shown in Figure 4. For 120 galaxies the stellar mass log M_* is between
9.45 and 10.03, and there after there are fewer galaxies up until the upper
mass range of 11.0. Please provide a Figure with the histogram of stellar
masses of the galaxies studied here. This sample is skewed towards galaxies
with lower stellar mass: a quick analysis of the galaxies in Kennicutt (1981)
which are also in S4G shows very few galaxies in the range of log M_* between
9.45 and 10.03. Please indicate what the consequences are of this selection
effect for the general conclusions.

Histogram added (hoping Tim has done this).

Since a sample of 98 galaxies were shown twice to (different?) volunteers, it
is of interest to compare the results for the same galaxies: did the resulting
pitch angles, number of arms, etc. all agree ? One of the data files available
on the GitHub website indicates that this is not the case, but that website is
poorly documented.

TODO: Compare spiral arms for validation subset

From Figure 10 of the GBpaper, it follows that 16 out of 98 galaxies have
p_bar .ge. 0.5, of which 8 out 98 galaxies have p_bar .ge. 0.55, with none of
them exceeding p_bar = 0.75. Thus there are few really barred galaxies in this
sample. This statistics can be repeated for the other sample of 100 galaxies.
This information is necessary to be able to judge how many barred galaxies are
considered. If only 16% of the galaxies are barred, there is not much you can
do to test hypotheses concerning barred galaxies.
 We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). No changes made (or asked for).

3. Section 3.1

You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than
2 arms per galaxy. Please give details of why this is so. Please also list how
many of the galaxies are multiarmed, and state the number of arms.

Tim working on this.

In Figure 4 I count about 40 1-armed spirals, 22 3-armed spirals, 14 4-armed
spirals, and about 63 2-armed spirals (more difficult to count). Why are there
so many 1-armed spirals, and why for these, for pitch angles below about 14
degrees, is PHI_arm not equal to PHI_gal ?

Tim says: Probably due to Aggregation. **Need to check**

4. Section 3.2.1

I quote: "Morphological classification commonly links bulge size to spiral
tightness, and such a link is implied by the Hubble Sequence (Sandage 2005,
Gadotti 2009, Buta 2013). Some studies have indeed reported a link between
measured spiral galaxy pitch angle and bulge size (i.e. Hart et al. 2017,
Davis et al. 2019), while others have not found any significant correlation
(Masters et al. 2019).

Now there is a large overlap in authors between Hart et al. (2017) and Masters
et al. (2019): 7 people, 5 of which are also co-authors of this paper. Surely
these 5 people should be able to explain WHY they found no link between pitch
angle and bulge size in one paper, and found a link in another one ??
Answering this is not in the scope of this paper.

Note to point out this is one of the ethical issues with this report.

Sandage (2005) reiterates that there are Sa galaxies with small bulges, so he
knew that there was a tension between the pitch angle criterion and the one
concerning the importance of the bulge. This is well explained in his 1961
Hubble Atlas of Galaxies. Please take this into account in your explanations.

We have added a citation.

5. Section 3.2.2

Here it is asserted that "To investigate this relationship in our data, we make
use of Galaxy Zoo 2's bar fraction (pbar), which has been demonstrated to be a
good measure of bar length (Willett et al. 2013) and bar strength (Skibba et
al. 2012; Masters et al. 2012; Kruk et al. 2018) and therefore a good measure
of the torque applied on the disc gas." In Athanassoula et al. (2009, MNRAS
394, 67) the bar strength is evaluated as the torque at the Lagrangian point:
Q_t,L1. Their figure 7 shows a sketch of the parameter R_D, which is the ratio
of the radius of the orbits in the crest in the outer ring to the radius of the
curve of zero velocity. A proxy could have been constructed from this Figure
for a measurement by citizen scientists by having them measure the radius of
the outer (pseudo-) ring to the radius of the L_4 point in the dark "banana"
part of the image. In any case in their figure 8 a correlation is found between
R_D and the bar strength. This figure is also shown in a different way as
Figure 5 of Athanassoula et al. (2009, MNRAS 400, 1706), the paper which the
authors cite.

Nothing asked for?

The manifold theory could be tested if Q_t,L1 can be suitably either mesured or
replaced by an appropriate proxy. As R_D can not be used (see discussion above)
the authors argue that Galaxy Zoo 2's bar fraction (pbar) is a good measure of
the bar strength. This, however, is not sufficient for the problem in question,
because not any measure of the bar strength may do, as Athanassoula et al. (2009,
MNRAS, 400, 1706) show in their figures 7 and 8, where a correlation is shown
with Qt,L1, but is practically lost if Qb is used instead. This could be because
Qt,L1 contains information on the bar strength, as well as on the bar pattern
speed which is also linked to the bar strength. Since pbar presumably does not
include information on the pattern speed it is possibly a better proxy for Qb
than for Qt,L1. This, however, together with a number of other points discussed
above remains to be shown. The authors could have consulted the volunteers with
in a more appropriate way so as to obtain answers for the question under
consideration. But, as it is, section 3.2.2 can not reach any conclusion. Thus,
either the whole of 3.2.2 should be omitted, or the last paragraph
should be reworded to explain that no conclusion could be reached.

We are reporting what we have asked the volunteers. We thank the referee for these suggestions for any potential future version of the site.  We can either make this more explicit, or try and find a better measurement. This is fair, and we should not be making specific claims regarding Manifold theory (it is still valid that we find no link between pitch angle and our measure of bar strength, which should be reported).

6. Section 3.3.3

Figure 8 has only one panel, so the sentences referring to the upper and lower
panels of this figure should be rephrased.

Corrected.

The range of values of PHI_gal is from 8 to about 50 degrees, and there is no
indication that the lower limit in the pitch angle is poorly determined. Note
that Kennicutt (1981) find a lower limit of 0 degrees. Setting the lower limit
to 15 degrees, so as just to be able to conclude that a uniform cot PHI_gal or
cot PHI_arm distribution cannot be excluded is thus completely arbitrary. It
would be best to display in a second panel the distribution of cot PHI_gal or
cot PHI_arm with a lower limit of 8 degrees, so as to alert the reader to the
strong influence of this assumption.

We believe the supporting text is sufficient to explain this. And have added some directly to the caption.

Obviously, the abstract and the conclusions should be toned down as well,
since a hurried reader would only retain your (possibly erroneous) conclusion.

We have toned down the abstract and conclusions.

7. Section 4

I quote: "We do not find any link between bar strength and pitch angle suggests
that the primary mechanism driving the evolution of the spirals in our sample
is not Manifold theory (see Section 3.2.2). Just in case you reach a conclusion
in that section, and want to mention it in this conclusion section, could you
please formulate this in proper English? With so many native English/American
speakers on the list of authors this ought not to be too much asked, and would
help the reader understand better, as well as save considerable work for the
copy editor.

We have reworded this section. 
This seems unnecessarily rude.

Another quote: "In this work, we assume that spiral arms are equally likely to
be identified and recovered at all pitch angles." OK, but then you should NOT
arbitrarely discard low pitch angle values when testing the Pringle & Dobbs
2019 hypothesis.

We assume no selection bias in galaxy builder results, which is not necessarily valid, however the requirement for a lower limit of pitch angle is needed for the test.

The last two paragraphs contain more advertisements for the work, rather than
Conclusions.

We have reworded this section.

8. Data availibility

The GitHub references refer to poorly documented working directories, and an
interested researcher needs to make a considerable effort to start finding the
basic information. It would be much more helpful to provide the most relevant
information about the only several 100s of galaxies in machine readable tables,
which can be stored at e.g. the CDS.

Tim TODO - checking with him.

To sum up: There are a number of problems with this paper, and a major revision
should be made, so as to rescue the valuable results provided by the volunteers,
and present them in a proper context.

We are happy the referee is convinced our experimental design is sound, and are positive about the annotations obtained. We also agree it needs to be published and presented to fully make use of the valuable results the volunteers provided.