### 2021-04-01 referee_response2.md
More tweaks on sample selection comments

# Referee Response, second round.

We thank the referee for these additional comments. As before, major changes have been hi-lighted in bold.

> Major comments:

> 1. In the Introduction, the authors change the three criteria of the ordering in the Hubble classification sequence. However, Sandage (2005) repeats these several times: the size of the central bulge, the degree of resolution into condensations, and the openness of the spiral arms.

> The authors change the "degree of resolution into condensations" to "the degree of resolution of the spiral arms (how 'patchy' they are)". Yet the word patchy is never used in this context, and for flocculent galaxies (where there are no well described spiral arms), such as NGC 2841 (type SAT3 in the Revised Shapley Ames Catalog), an ordering is nonetheless given, in coherence with all three criteria.

> Please refrain from reformulating the Hubble classification to a scheme based on your own criteria: e.g. the prototype flocculent spiral galaxy NGC 2841 has condensations in its disc, but no clear spiral arms.

We have made edits to use the recommended language.

> 2. In this context, it is striking that the authors now choose NGC 4689 as a example of a flocculent galaxy, even though spiral structure is visible, in particular the eastern arm. True, Elmegreen & Elmegreen (1987) list it as a class 3 spiral, i.e. in the same category as NGC 2841, but the spiral arms are clearly there in the 3.6 micron Spitzer image. I suggest using NGC 2841 as a good example for a flocculent galaxy, and note that for NGC 5055 there is a paper by Michele Thornley -  1996ApJ...469L..45T - pointing out that in the K'-band there is spiral structure seen, even though it is also listed as a prototype flocculent spiral in Elmegreen & Elmegreen (1987).

We have edited to use NGC 2841 as the example of a flocculent galaxy ** TO DO - I put the SDSS image of it in the drive.  **

> 3. The authors write on page 2 left side line 38: Spiral tightness is also observed to correlate with rotation curve shape (Seigar et al. 2005), with rising rotation curves creating more open spiral structure.

> I think you mean that galaxies with rising rotation curves have more open spiral structure. I don't think rotation curves are creating anything. More likely, the associated lower shear allows more open spirals.

We have edited this sentence as suggested.

> 4. In section 3.2.2 the authors write: One of the predictions of Manifold theory is that pitch angle increases with bar strength (Athanassoula et al. 2009b).

> Please make this section self-contained: ... pitch angle increases with bar strength (Athanassoula et al. 2009b), as evaluated at the Lagragian L_1 and L_2 points. The bar strength evaluated that manner is very specific, and in Athanassoula et al. 2009b it is shown that a correlation with Q_bar is much less strong. Since you use here p_bar as a proxy for bar strength, it is only remotely correlated with the bar strength defined in Athanassoula et al. 2009b. True, you admit that your test is not very convincing, but the interested reader should not have to consult explicitly the Athanassoula et al. 2009b paper to find out why.

We have edited this section as suggested.

> 5. In the conclusion section, the formulation of the conclusions on section 3.2.2. are much stronger than justified: you say that you cannot completely rule out the manifold theory, even though you admit that the proxy you use is not convincing. Since hurried readers could read this conclusion section without reading the bulk of the paper, you leave them with the impression that you almost rule out the Manifold theory, while this is not the case, since your proxy for bar strength is very poor, and your sample is small. Please harmonize your conclusions with the reality you describe in the main section.

We have further toned down the conclusion on Manifold theory and believe it now includes most of the caveats given in the main section.

> 6. In the conclusions, page 11 left side lines 48-49, you suggest that the absence of galaxies with low pitch angles is not a selection effect. How does this square with the data of Kennicutt (1981), where nearly all Sa galaxies, as classified in the RSA catalogue, have pitch angles less than 10 degrees? Are you sure that you are not missing the tightly wound early type spiral galaxies?

Our sample selection is quite different to Kennicutt (1981) which may be one factor. For example the most distant of the Kennicutt galaxies is at a similar redshift the lower limit of our sample. This means the mass range covered will be very different too.  It seems possibly that very tightly wound spirals may be more likely to be misidentified as smooth galaxies, and therefore not make it into the spiral sample in Galaxy Zoo at all. In addition a colleague compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals. We have not run the Kennicutt sample through Galaxy Builder, so we cannot reliably comment if that would be an effect here, but will make sure to do that for the next phase of the project.

> I did a cursory inspection of the early type spirals in the Yu & Ho (2018) sample, and find that most of the early type spirals (S0/a, Sa and Sab) are actually barred spirals when checking the classification in Hyperleda and displaying them in the legacysurvey viewer. Perhaps the authors can check the distribution of the results of Kennicutt (1981) in a diagram such as Figure 9, to see whether his sample is in agreement with the ideas of Pringle & Dobbs (2019)?

> Minor comments and typos:

Thank you for noting these minor typos and other errors. We have corrected these, and a handful of other typos we noted in editing.

> 1. page 2 right side line 50: disk - please use English spelling throughout: disc.
> 2. page 3 left side line 40: we further separate the galaxies ...
> 3. page 4 left side line 21: and report the pitch angle
> 4. page 4 left side line 53: which results in a sample favouring (3 times "which" is too much)
> 5. page 4 left side line 60: of these classifications
> 6. page 6 right side line 19: morphological
> 7. page 10 right side line 27: you write "pitch angle calculation; obtaining" but the ; should disappear